---
title: "Course Material [2]"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

## 5. Import and Export Data

The example dataset steam_subset.csv can be downloaded [here](https://gofile.io/?c=7bP3y4). The colums are seperated with comma( , ) and the first line is the column names.

Read the data set from a file with `read.table`, use functions `head` and `str` to check the dataset

```{r}
steam<- read.table(file = "steam_subset.csv", sep = ",", header = T)
head(steam)
str(steam)
```


or use `read.csv` to read the file

```{r}
steam<- read.csv(file = "steam_subset.csv", header = T, sep = ",")
head(steam)
str(steam)
```

Computation on the variables
```{r}
mean(steam$Level)
sd(steam$Level)
```

Attach a dataset
```{r}
attach(steam)
mean(Level)
sd(Level)
```



## 6. Statistical Computation

### Distributions

* `d`xxx: density
* `q`xxx: quantile
* `p`xxx: cumulative 
* `r`xxx: randome samples

Take normal distribution for an example

```{r}
dnorm(0, 0, 1)
qnorm(.975, 0, 1)
pnorm(1.96, 0, 1)
rnorm(10, 0, 1)
```

### Student t-test
```{r}
```

### Analysis of Variance (ANOVA)
```{r}
```

### Linear Regression
```{r}
```

## 7. Plot and Graphics

* `save.image()` 

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## 8. Packages