---
title: "Course Material [2]"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

## 5. Plot and Graphics

Some commonly used functions

* `plot()` 
* `points()` 
* `lines()`
* `test()`
* `mtest()`
* `abline()`

```{r}
x<- 1:5
y<- c(3, 2, 4, 1, 5)
plot(x, y, type = "p", main = "point")
plot(y, type = "p", main = "point")
plot(x, y, type = "l", main = "line")
plot(x, y, type = "o", main = "line and point")
```

Add points, lines, title and legend
```{r}
x<- seq(0, 2*pi, .2)
plot(x, sin(x), type = "o", pch = 19)
points(x, cos(x), pch = 19, col = 2)
lines(x, cos(x), lty = 2, col = 2)
legend(4, 1, legend = c("sin", "cos"), pch = c(1, 19), lty = c(1, 2), col = c(1, 2))
```

`pch`: point style
```{r}
plot(sapply(1:5, function(x) rep(x, 5)), rep(5:1, 5),
     pch = 1:25, xlab = "", ylab = "", main = "", xlim = c(0.7, 5.2), axes = F)

text(sapply(1:5, function(x) rep(x, 5)) -.2, rep(5:1, 5), 1:25)
```

`lty`: line style

```{r}
plot(NULL, xlim = c(0,4), ylim = c(1,6), xlab = "", ylab = "", axes = F)
for(i in 1:6){
  y_pos<- 6:1
  lines(c(1, 4), c(y_pos[i], y_pos[i]), lty = i)
  text(0.5, i, paste("lty = ", y_pos[i]))
}
```

`color`: color
```{r}
plot(sapply(1:5, function(x) rep(x, 5)), rep(5:1, 5),
     pch = 16, col = 1:25, xlab = "", ylab = "", main = "", xlim = c(0.7, 5.2), axes = F)

text(sapply(1:5, function(x) rep(x, 5)) -.2, rep(5:1, 5), 1:25)
```

### Histogram
```{r}
attach(iris)
hist(Sepal.Length)
```

### Barplot
```{r}
attach(chickwts)
barplot(table(feed))
```

### Boxplot
```{r}
boxplot(weight)
```

Group-wise boxplot
```{r}
boxplot(weight~feed)
```

### Multiple plots
```{r}
par(mfrow = c(2,2))
barplot(table(feed))
hist(weight)
boxplot(weight)
boxplot(weight~feed)
```

```{r}
attach(chickwts)
layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
boxplot(weight~feed)
hist(weight)
boxplot(weight)
```

### Scatter plot for more than 2 variables
```{r}
pairs(trees)
```

### Plot into a file
```{r, eval = F}
png(file = "exmple.png", bg = "transparent")
plot(x, sin(x), type = "o")
dev.off()
```

### *Exercise 5* 
(attach the corresponding plots to the email)

1. Use the dataset `airquality`. Draw a scattar to show the relationship between variables `Ozone` and `Solar.R`. Show the points with `Wind` values greater than 9.7 in red, otherwise in black.

2. Make plot to monitor the convergence of the exponential function ($e^x \ = \sum_{n=0}^{\infty} \frac{x^{n}}{n!}$) with $x = 5$. Save the plot and attach to the email.

```{r, eval = F}
taylor_appox<- function(x){
  function(N){
    # TODO
  }
}

x<- 5
taylor_appox_N<- taylor_appox(x)

N_vec<- 1:20
approx_val<- sapply(N_vec, function(N){
  # TODO
}} 
plot(N_vec, approx_val)
```

3. Use the dataset `sleep`. Put a histogram and a boxplot in one plot to show the overall pattern of the increased sleeping time (variable `extra`).

4. Use the dataset `sleep`. Use boxplots to show the difference of the increased sleeping time between two groups.

5. Use the vector `islands`. Use barplots to disply the area of the 2 smallest islands and the 2 biggest islands. (Hint. you can use `sort(, decreasing = )` to sort a vector.)

6. Use dataset `iris`. Make a scatter plot of four variables `Sepal.Length`, `Sepal.Width`, `Petal.Length` and `Petal.Width`. Show different species in different colors. (e.g. `setosa` in black, `versicolor` in red and `virginica` in green)

7. Use dataset `ChickWeight`. From each of the `Diet` group, select 1 chick (identifier variable `Chick`). Make a time series plot to show the weight changing over time of the 4 selected chicks, each in one different color. Put a legend on the plot which contains the information of the selected chicks (`Diet` and `Chick`) and their corresponding colors.

## 6. Statistical Computation

### Descriptive Statistics

* `mean()`: mean
* `var()`: variance
* `sd()`: standard deviation
* `median()`: median
* `quantile()`: sample quantiles
* `min()`: minimum
* `max()`: maximum
* `range()`: range (minimum and maximum values)
* `summary()`: summary
* `IQR()`: interquartile range
* `cov()`: covariance
* `cor()`: correlation

```{r}
quantile(trees$Height, prob = c(.05, .15, .25, .75))

mean(trees$Height)

var(trees$Height)

cov(trees)

cor(trees)

summary(trees)

apply(trees, 2, range)

apply(trees, 2, sd)

apply(trees, 2, IQR)
```

### Distributions

* `d`xxx: density
* `q`xxx: quantile
* `p`xxx: cumulative 
* `r`xxx: randome samples

E.g. Normal distribution

```{r}
dnorm(0, 0, 1)
qnorm(.975, 0, 1)
pnorm(1.96, 0, 1)
rnorm(10, 0, 1)

x<- rnorm(100000, 0, 1)
hist(x, freq = F, ylim = c(0, .5))
lines(density(x))
```

`density`: kernel density estimation.

E.g. Beta distribution

```{r}
dbeta(0.5, 1, 2)
qbeta(.975, 1, 2)
pbeta(0.5, 1, 2)
rbeta(10, 1, 2)
```

E.g. Gamma distribution

```{r}
dgamma(10, 5)
qgamma(.5, 5)
pgamma(10, 5)
rgamma(10, 5)
```

Other commonly used distributions 

* Uniform (`dunif`, `qunif`, `punif`, `runif`)
* Binomial (`dbinom`, `qbinom`, `pbinom`, `rbinom`)
* Poisson (`dpois`, `qpois`, `ppois`, `rpois`)
* Cauchy (`dcauchy`, `qcauchy`, `pcauchy`, `rcauchy`)
* Student t (`dt`, `qt`, `pt`, `rt`)
* Weibull (`dweibul`, `qweibul`, `pweibul`, `rweibul`)
* Geometric (`dgeom`, `qgeom`, `pgeom`, `rgeom`)
* ...

### Student's *t*-test
```{r}
attach(iris)
levels(Species)
t.test(Sepal.Length[Species == "setosa"], Sepal.Length[Species == "versicolor"])
```

### Analysis of Variance (ANOVA)
```{r}
aov_res<- aov(Sepal.Length ~ Species)
summary(aov_res)
```

### Linear Regression
```{r}
attach(airquality)
lm_res<- lm(Ozone ~ Solar.R + Wind + Temp)
summary(lm_res)

```

Model selction via stepwise approach

```{r}
lm_step<- step(lm_res)
summary(lm_step)
```

Prediction
```{r}
input<- data.frame("Solar.R" = c(10, 100), "Wind" = c(20, 5), "Temp" = c(76, 76))
predict(lm_step, input)
```

### *Exercise 6*
[To be announced]

1. Use the dataset `iris`, calculate the corrlations among variables `Sepal.Length`, `Sepal.Width`, `Petal.Length` and `Petal.Width`.

2. Use the dataset `iris`, calculate the summary statistics of variables `Sepal.Length`, `Sepal.Width`, `Petal.Length` and `Petal.Width`.

3. Exploring _Centrial Limit Theorem_. Generate 100 samples from a poisson distribution (`rpois`) with parameter $\lambda = 2$, calculate the mean value out of the 100 samples. Repeate the process 10000 times. Make a histogram of the 10000 mean values. Check the empirical mean and the standard deviation of the 10000 mean values, compare to the theoretical values $\mu = 2$ and $\sigma = \frac{\sqrt{2}}{10}$.

4. Use the dataset `sleep`, conduct a statistical test of the hypothesis that there is no difference of the increased slepping hours between two groups. The significant level $\alpha = 0.05$. (Hint: `paired = T`).

5. Use the dataset `chickwts`, conduct a statistical test of the hypothesis that the feed type doesn't affect the weight of chicks.

6. Use the dataset `cars`, build a linear model to examine how the speed (`speed`) affacts the stopping distance (`dist`).

## 7. Packages
The packages such as `stats`, `base` are autimatically loaded.

```{r}
(.packages())
```

You can use `library` to load packages which are already installed.

Example: Single-Layer Neurual Network
```{r}
library(nnet)

## split data into training and testing datasets
n_samples<- nrow(iris)
train_ind<- sample(1:n_samples, n_samples * (4/5))
test_ind<- (1:n_samples)[-train_ind]

train_dat<- iris[train_ind,]
test_dat<- iris[test_ind,]

nn_model<- nnet(Species ~., linout = T, size = 5, data = train_dat)

pred_nn<- predict(nn_model, test_dat, type = 'class')

## Accuracy
table(test_dat$Species == pred_nn) / nrow(test_dat)
```

Example: Decision Tree
```{r}
library(rpart)

cart_model<- rpart(Species ~., data = test_dat)

pred_cart<- predict(cart_model, test_dat, type = 'class')

## Accuracy
table(test_dat$Species == pred_cart) / nrow(test_dat)
```

Use `install.packages()` to install package.

Example: Use t-SNE for dimensionality reduction, visualizing breast cancer data.

```{r}
library(tsne)

url<- "https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/wdbc.data"
exp_dat<- read.table(url, sep = ",", header = F)

tsne_res<- tsne(exp_dat[, 3:32], perplexity = 100)

plot(tsne_res[which(exp_dat$V2 == "B"), ], col = 1, pch = 20, xlim = c(-6, 6), ylim = c(-6, 6), xlab = "", ylab = "")
points(tsne_res[which(exp_dat$V2 == "M"), ], col = 2, pch = 20)
```

<!-- `installed.packages`, `install.packages`, `(.packages())`, `library` -->