---
title: "Introduction to R [1]"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```

This is a tutorial document for the Tampere Summer school course **Introduction to R** made by the course lecturer [Chien Lu](https://humblelu.github.io/personal_website/). You can find it from the corresponding course  [webpage](https://www.tuni.fi/en/study-with-us/introduction-r-tampere-summer-school).

## Course Logistics

* Leaning outcome: After this course, participants will be able to use R to conduct basic statistical analysis and data visualization
* Assessment: Pass/Fail
* Course Requirement:
    + Deadline: 31.08
    + 6 Exercise packs 
    + Data analysis course work (individual or group)
        + Perform statistical data analysis using R
        + If you don't have your own dataset, come to discuss with me!
    + No exams

## Learning R?

* Free (impotrant!)
* Massive statistical packages
* More and more popular in both academia and industry
* _... Millions of  other reasons_

### Further Reading

* Matloff, Norman. _The art of R programming: A tour of statistical software design_. No Starch Press, 2011.

## Before Starting

* Download R from [CRAN](https://cran.r-project.org/mirrors.html)
* You can also Download an useful GUI called [RStudio](https://www.rstudio.com/products/rstudio/)

## Getting Started!

### RStudio
Create a script by clicking _File_ >> _New File_ >> _R Script_

To execute your code, highlight the code you wish to execute and press _Ctrl + Enter_

Show the current working directory

### RGui (Windows)
Create a script by clicking _File_ >> _New Script_

To execute your code, highlight the code you wish to execute and press _F5_

### RGui (Mac OS)
Create a script by clicking _File_ >> _New Document_

To execute your code, highlight the code you wish to execute and press _Ctrl + Enter_

### Set up the working directory

```{r, eval=FALSE}
getwd()
```

List all the files under the current working directory

```{r, eval=FALSE}
dir()
```

Change the working directory (mind the `/`)

```{r, eval=FALSE}
setwd("C:/Users/cl422876/Documents")
```

### Quit the R session

```{r, eval=FALSE}
q()
```

## R as Calculator

You can start with playing around with R, use it as a calculator

```{r}
1 + 3
6*(5-1)
```

you can also try out some more complicated (fancier) computations, such as Trigonometric functions (high school nightmare):

* sin($\frac{\pi}{2}$) = ?
* cos($\pi$) = ?
* tan($\frac{\pi}{4}$) = ?

find out the answers using functions `sin()`, `cos()`, and `tan()`, for example:

```{r}
sin(pi/2)
```

or other mathematical functions are available, such as: 

* `exp()`: Exponential with base _e_
* `log()`: Logarithm
* `log10()`: Logarithm with the base 10
* `sqrt()`: Square root
* `abs()`: Absolute value
* `round()`: Round the value
* `floor()`: Round down the value
* `ceiling()`: Round up the value
* `factorial()`: Factorial function

Verify if the function `factorial` returns the correct value
```{r}
factorial(5)
5*4*3*2*1
```

You also can use `function` to define your own function 

```{r}
my_square<- function(x){
  x^2
}

my_square(4)
```

where `<-` is the assignment operator

<!-- or scientific notations -->

<!-- ```{r} -->
<!-- 9.6e-4 -->
<!-- ``` -->

## Basic R Objects

### Assign value to an

Use `<-` to assign a value to an object
```{r}
x<- 1
x
```

of course, `=` also works
```{r}
x = 1
x
```

### Numbers

```{r}
x<- 7	                # integer
y<- 0.8	              # numeric (real number)
z<- 3+0i            	# complex
class(z)              # gives the class of z
x<- 9.6e-4            # scientific notation

Inf		                # infinity
1/Inf
Inf/1
NaN       	          # not a number
NA	                	# not available/missing value

is.na(x)	            # tests whether x is  "not available" or not (=available)
is.nan(x)             # tests whether x is "not a number" or not
```


### Logical Operations


### Vectors

All the elements in a vector should be of the same object type. 

Use `c` to create a *vector*.
```{r}
## number
exp_1<- c(1, 2, 3, 4, 5)
exp_1

## logical
exp_2<- c(TRUE, FALSE, FALSE, TRUE)
exp_2

## string
exp_3<- c("I", "am", "a", "meaningless", "example")
exp_3
```

Use `seq` to create a vector with sequential numbers
```{r}
a<- seq(from = 1, to = 5, by = 1)
a
```

or just simply:
```{r}
a<- 1:5
a
```

* `min()` and `max()`: Minimum value and maximum value within a vector
* `which.min()` and `which.max()`: Index of the minimal element and maximal
element of a vector
* `pmin()` and `pmax()`: Element-wise minima and maxima of several vectors
* `sum()` and `prod()`: Sum and product of the elements of a vector
* `cumsum()` and `cumprod()`: Cumulative sum and product of the elements of a
vector

Concatenate vectors
```{r}
vec_1<- c(1, 1, 1)
vec_2<- c(2, 2, 2)
vec_join<- c(vec_1, vec_2)
vec_join

vec_3<- c(3, 3, 3)
vec_join<- c(vec_1, vec_2, vec_3)
vec_join
```

Subset a vector
```{r}
a<- c(1, 2, 3, 4, 5)

# extract with indices
a[c(1, 3, 5)]

# extract with logicals
a[c(T, F, T, F, T)]

a %% 2 == 1

a[(a %% 2 == 1)]

# omit
a[-c(2, 4)]

a[-which(a %% 2 == 0)]

```

`NA` values in a vector
```{r}
a<- c(1, NA, 2, NA, 3)
a

b<- c(1, 2, 3, 4, 5)
b * c(1, NA, 1, NA, 1)

# replace NA with 0
a[is.na(a)]<- 0
a
```

### Strings
```{r}
a <- "hello"          # we can have objects that are strings
a
class(a)              # or typeof(a): gives the class of the object a
print("Hello R!")     # prints thw text in the argument of the function
```
### Factors



### Matrices

## Exercise Pack 1

1. Set the working directory to your Desktop and use `getwd` to verify
2. Create a vector which its variance value equals to 0 and mean value equals to 5.
3. Create a function to convert the degree Celsius to degree Fahrenheit
4. Create a function which returns the sum of the maximum and the mininum value of the input vector

Use `rep` to create a vector with replicate elements
```
b<- rep(x = 1, times = 3)
b
```


For loop
```{r}
for(i in 1:5){
  print(i)
}
```

While loop
```{r}
i<- 1
while(i < 5){
  i<- i + 2
}
```

Apply functions

## Exercise Pack

1. Using a loop two print out all the prime numbers smaller than 50
2. 
3. 


### Useful tips

You can use `#` to add comment, the code after will not be executed.

```{r}
a<- c(1, 2, 3) #I have no idea why I created this vector
```

Use `?` to call R help when you have difficulties.

```{r}
?c
```

For efficient computation, try to use apply functions instead of for-loop

## Import and Export Data

Read the data set from a file with `read.table()`

```{r}
```



## Packages

## Statistical Computation

### Basic Operations

```{r}
attach(iris)
mean(Sepal.Length)
sd(Sepal.Length)
median(Sepal.Length)
```

### t-test

### Analysis of Variance (ANOVA)

### Distributions

* `d`xxx: density
* `q`xxx: quantile
* `p`xxx: cumulative 
* `r`xxx: randome samples

Take normal distribution for an example

```{r}
dnorm(0, 0, 1)
qnorm(.975, 0, 1)
pnorm(1.96, 0, 1)
rnorm(10, 0, 1)
```

## Plot and Graphics

* `save.image()` 

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
